Decoupled wood and leaf strategies among canopy and understorey tree in a low-elevation rainforest of India

### 2018-02-13 inst/doc/draft2.Rmd
13/02/18 Draft

---
title: Decoupled wood and leaf strategies among canopy and understorey tree in a low-elevation rainforest of India
author: "Sylvain Schmitt sylvain.schmitt@agroparistech.fr"
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
  bookdown::pdf_document2:
    number_sections: false
csl: mee.csl
bibliography: refs.bib
link-citations: yes
---

@@

```{r Traits}
PFT <- genPFT()
kable(data.frame(
  Trait = c('Leaf thickness', 'Leaf area', 'Leaf dry matter content', 'Specific leaf area', 'Wood density'),
  Abbreviation = c('LT', 'LA', 'LDMC', 'SLA', 'WD'),
  Unit = c('$\\mu m$', '$mm^2$', '$mg.g^-1$', '$m^2.kg^-1$', '$g.cm^-3$'),
  Strategy = c('Leaf defense and investment',
        'Leaf resource capture vs.investment',
        'Leaf defense and investment',
        'Leaf resource acquisiton vs. defense',
        'Stem transport, structure and defense')
), caption = 'Functional traits measured. Traits with their abbreviation, standard unit, and associated species strategy for functioning, survival and reproduction [@baraloto_decoupled_2010].')
```

@@

We calculated weighted mean (GWM, where G stands for "guild") and weighted variance (GWV) [GWM and GWV, @violle_let_2007] for each guild in each community. In the quadrat $j$, GWM of trait $t$ was the mean of trait values $t_{i}$ of species $i$ weighted by species relative abundance $p_{i}$:
$$GWM_{j} = \sum_{i=1}^{S} p_{i,j} * t_{i,j} ~ (1)$$
where $S$ is the number of species in the quadrat $j$. We also calculated the average trait values based on presence-absence values to assess the influence of less abundant and rare species on community trait composition [@cingolani_filtering_2007]. GWV of trait $t$ was calculated as :
$$GWV_{j} = \sum_{i=1}^{S} p_{i,j} * (t_{i,j} - GWM_{j})^2 ~ (2)$$

To test for any significant changes in functional composition, we devised a null model approach. We thus explored following questions with subsequent hypothesis:

* __Q1__ Does local functional composition from guilds differ ?
    * ___H1a__ Environmental filtering differ in selection of local functional composition between canopy and understorey species._
    * ___H1b__ Environmental filtering strenght differ locally between canopy and understorey species._
* __Q2__ Does the local functional composition reflect the influence of abiotic and biotic drivers of community dynamics within guilds ?
    * ___H2a__ Drivers of environmental filtering on local functional composition differ between canopy and understorey tree species._ 
    * ___H2b__ Strength of drivers of evironmental filtering differ between canopy and understorey tree species._

To accept or reject hypothesis we used three null models :

* __NM1__ randomized species among guilds. Consequently, NM1 breaks the link between species and guilds while keeping community structure, thus testing for hypothesis __H1a__.
* __NM2__ randomized trait values among species in the whole plot for each guild (permuting raw in `Trait*Species` matrices of both guilds) [@bernard-verdier_community_2012]. Consequently, NM2 breaks the link between environement and local functional composition while keeping guilds and species spatial structure, thus testing for hypothesis __H2a__.
* __NM3__  randomized abundances among species within each guild but kept the observed list of species and trait values unchanged inside guilds [@bernard-verdier_community_2012]. Consequently NM3, breaks the link between environment and species abundances while keeping guilds, community, and species spatial structure, thus testing for hypothesis __H1b__ and __H2b__.

@@

To characterize the patterns of deviation from the null model, we also calculated standardized effect sizes (SES) as
$$SES = \frac{(I_{obs} - I_{null})}{\sigma _{null}} (2)$$
where $I_{obs}$ was the observed metric and $I_{null}$ and $\sigma _{null}$ were the mean and standard deviation of the corresponding null distribution. The influence of environment on SES values was investigated by linear regression with biotic and abiotic factors ($SES_{trait} = Slope + Curvature + Wetness + SW + Canopy + BA$). Analyses were conducted on the five functional traits.

@@

```{r CV}
PFT %>% 
  select(-Tree, -Sp, -SpCode, -CE, -SHA, -LTD) %>% 
  melt(variable.name = 'trait') %>% 
  group_by(trait) %>% 
  summarise_each(funs(min(.,na.rm = T), max(., na.rm = T), 
                      mean(.,na.rm = T), sd(.,na.rm = T))) %>% 
  left_join(PFT %>% 
              select(-Sp, -SpCode, -CE, -SHA, -LTD) %>% 
              melt(id.vars = 'Tree', 
                   variable.name = 'trait', 
                   value.name = 'Intra-individuals') %>% 
              group_by(Tree, trait) %>% 
              summarise_each(funs(sd(., na.rm = T)/mean(., na.rm = T))) %>% 
              ungroup() %>% 
              select(-Tree) %>% 
              group_by(trait) %>% 
              summarise_each(funs(mean(., na.rm = T)))) %>% 
  left_join(PFT %>% 
              group_by(Tree, Sp, SpCode, CE) %>% 
              summarise_each(funs(mean)) %>% 
              ungroup() %>% 
              select(-Tree, -SpCode, -CE, -SHA, -LTD) %>% 
              melt(id.vars = 'Sp', 
                   variable.name = 'trait', value.name = 'Intra-species') %>% 
              group_by(Sp, trait) %>% 
              summarise_each(funs(sd(., na.rm = T)/mean(., na.rm = T))) %>% 
              ungroup() %>% 
              select(-Sp) %>% 
              group_by(trait) %>% 
              summarise_each(funs(mean(., na.rm = T)))) %>% 
  left_join(PFT %>% 
              group_by(Tree, Sp, SpCode, CE) %>% 
              summarise_each(funs(mean)) %>% 
              ungroup() %>% 
              select(-Tree, -CE) %>% 
              group_by(Sp, SpCode) %>% 
              summarise_each(funs(mean)) %>% 
              ungroup() %>% 
              select(-Sp, -SpCode, -SHA, -LTD) %>% 
              melt(variable.name = 'trait', value.name = 'Inter-species') %>% 
              group_by(trait) %>% 
              summarise_each(funs(sd(., na.rm = T)/mean(., na.rm = T)))) %>% 
  kable(digits = 2, caption = 'Intra-individual, intra-specific and inter-specific component of trait variation. The overall range, mean and standard deviation are provided for each trait. The remaining columns represent the coefficient of variation of trait values among leaves per individuals (intra-individual), among trees per species (intra-specific), and among trees of different species. See table \\@ref(tab:Traits) for abbreviation.')
```

The first axis of the principal component  analysis (PCA) of mean species trait values was basically related to wood density and structural leaf trait variations (LT, LA, LDMC), thus opposing species with thick, big leaves and lightwood to species with dense leaves and hardwood. The second axis mostly opposed SLA to LDMC, and to a less extent to wood density. However, wood density was orthogonal to SLA in the plan 1-2 of the PCA indicating independence between those two traits. Finally, hypervolume overlap between the two guilds was not significantly lower than null expectations ($p = 0.165$, see Supplementary materials [S2: Hypervolumes of functional guilds]).

@@

#### H1a. Environmental filtering differ in selection of local functional composition between canopy and understorey species

```{r GWMH1a, fig.cap="Number of communities significantly (p-value<0.05) different from null expectations with null model NM1 for each trait of GWMs, for canopy (C) and understorey (U) guilds, in abundance. See table \\@ref(tab:Traits) for abbreviations.", fig.height=4}
GWM <- L %>%
  t() %>%
  data.frame() %>%
  mutate(SpCode = row.names(.)) %>%
  left_join(Q) %>%
  select(-SpCode) %>%
  melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
       variable.name = 'community', value.name = 'abundance') %>%
  group_by(strata, community) %>%
  summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
# nQ <- array(rep(as.matrix(Q), n),
#             c(dim(Q), n),
#             list(row.names(Q), names(Q), 1:n))
# nQ <- array(apply(nQ, 3, function(x){
#   x[,'strata'] <- sample(x[,'strata'])
#   return(x)}), c(dim(Q), n),
#   list(row.names(Q), names(Q), 1:n))
# cl <- makeCluster(cores)
# clusterExport(cl, list('L', 'nQ'))
# nGWM <- parApply(cl, nQ, 3, function(x){
#   library(dplyr)
#   library(reshape2)
#   L %>%
#     t() %>%
#     data.frame() %>%
#     mutate(SpCode = row.names(.)) %>%
#     left_join(data.frame(x)) %>%
#     select(-SpCode) %>%
#     melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
#          variable.name = 'community', value.name = 'abundance') %>%
#     group_by(strata, community) %>%
#     mutate_each(funs(as.character), -abundance) %>%
#     mutate_each(funs(as.numeric), -abundance) %>%
#     summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
# })
# stopCluster(cl)
# nGWM <- array(unlist(lapply(nGWM, function(x) as.matrix(x[3:9]))),
#               c(dim(GWM[3:9]), n),
#               list(row.names(GWM), names(GWM[3:9]), 1:n))
# rm(cl, nQ)
# save(nGWM, file = file.path(path, 'nGWM_NM1.Rdata'))
load(file.path(path, 'nGWM_NM1.Rdata'))
GWM_pval <- GWM
GWM_pval[3:9] <- apply(
  array(c(as.matrix(GWM[3:9]), nGWM), c(dim(GWM[3:9]), n+1)), 
  1:2, function(x) rank(x)[1]/(n+1))
GWM_pval %>% 
  select(-community, -LES, -WES) %>% 
  group_by(strata) %>% 
  mutate_all(function(x) findInterval(x, c(0.05, 0.95)) != 1) %>% 
  melt(id.vars = "strata", variable.name = "trait") %>% 
  filter(value == TRUE) %>% 
  ggplot(aes(x = factor(trait), 
             fill = factor(strata))) +
  geom_bar(position = "dodge") +
  scale_fill_discrete("Guild") +
  xlab("Functional trait") +
  ylab("Number of communities") +
  ylim(c(0,dim(L)[1]))
```

#### H1b. Environmental filtering differ in selection of local functional composition between canopy and understorey species

```{r GWVH1b, fig.cap="Number of communities significantly (p-value<0.05) different from null expectations with null model NM3 for each trait of GWVs, for canopy (C) and understorey (U) guilds, in abundance. See table \\@ref(tab:Traits) for abbreviations.", fig.height=4}
GWV <- L %>%
  t() %>%
  data.frame() %>%
  filter(sum(.) > 2) %>%
  mutate(SpCode = row.names(.)) %>%
  left_join(Q) %>%
  select(-SpCode) %>%
  melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
       variable.name = 'community', value.name = 'abundance') %>%
  melt(id.vars = c('strata', 'community', 'abundance'),
       variable.name = 'trait', value.name = 'trait_value') %>%
  left_join(melt(GWM, id.vars = c('strata', 'community'),
                 variable.name = 'trait', value.name = 'GWM')) %>%
  group_by(strata, community, trait) %>%
  mutate(GWV = (trait_value - GWM)^2) %>%
  select(-trait_value, -GWM) %>%
  summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) %>%
  dcast(strata + community ~ trait)
# nL <- array(replicate(n, t(apply(L, 1, function(x){
#   x[Q[which(Q$strata == "C"), "SpCode"]][x[Q[which(Q$strata == "C"), "SpCode"]] != 0] <- sample(x[Q[which(Q$strata == "C"), "SpCode"]][x[Q[which(Q$strata == "C"), "SpCode"]] != 0])
#   x[Q[which(Q$strata == "U"), "SpCode"]][x[Q[which(Q$strata == "U"), "SpCode"]] != 0] <- sample(x[Q[which(Q$strata == "U"), "SpCode"]][x[Q[which(Q$strata == "U"), "SpCode"]] != 0])
#   return(x)
#   }))), c(dim(L), n), list(row.names(L), names(L), 1:n))
# cl <- makeCluster(cores)
# clusterExport(cl, list('nL', 'Q', 'GWM'))
# nGWV <- parApply(cl, nL, 3, function(x){
#   library(dplyr)
#   library(reshape2)
#   x %>%
#     t() %>%
#     data.frame() %>%
#     filter(sum(.) > 2) %>%
#     mutate(SpCode = row.names(.)) %>%
#     left_join(Q) %>%
#     select(-SpCode) %>%
#     melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
#          variable.name = 'community', value.name = 'abundance') %>%
#     melt(id.vars = c('strata', 'community', 'abundance'),
#          variable.name = 'trait', value.name = 'trait_value') %>%
#     left_join(melt(GWM, id.vars = c('strata', 'community'),
#                    variable.name = 'trait', value.name = 'GWM')) %>%
#     group_by(strata, community, trait) %>%
#     mutate(GWV = (trait_value - GWM)^2) %>%
#     select(-trait_value, -GWM) %>%
#     summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) %>%
#     dcast(strata + community ~ trait)
# })
# stopCluster(cl)
# nGWV <- array(unlist(lapply(nGWV, function(x) as.matrix(x[3:9]))),
#               c(dim(GWV[3:9]), n),
#               list(row.names(GWV), names(GWV[3:9]), 1:n))
# rm(cl, nL)
# save(nGWV, file = file.path(path, 'nGWV_NM3.Rdata'))
load(file.path(path, 'nGWV_NM3.Rdata'))
GWV_pval <- GWV
GWV_pval[3:9] <- apply(
  array(c(as.matrix(GWV[3:9]), nGWV), c(dim(GWV[3:9]), n+1)), 
  1:2, function(x) rank(x)[1]/(n+1))
GWV_pval %>% 
  select(-community, -LES, -WES) %>% 
  group_by(strata) %>% 
  mutate_all(function(x) findInterval(x, c(0.05, 0.95)) != 1) %>% 
  melt(id.vars = "strata", variable.name = "trait") %>% 
  filter(value == TRUE) %>% 
  ggplot(aes(x = factor(trait), 
             fill = factor(strata))) +
  geom_bar(position = "dodge") +
  scale_fill_discrete("Guild") +
  xlab("Functional trait") +
  ylab("Number of communities") +
  ylim(c(0,dim(L)[1]))
```

@@

#### H2a. Drivers of environmental filtering on local functional composition differ between canopy and understorey tree species

>>>>

---
title: Decoupled wood and leaf strategies among canopy and understorey tree in a low-elevation rainforest of India
author: "Sylvain Schmitt sylvain.schmitt@agroparistech.fr"
date: '`r Sys.Date()`'
output:
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
linestretch: 1.5
csl: mee.csl
bibliography: refs.bib
link-citations: yes
---

@@

```{r Traits}
PFT <- genPFT()
kable(data.frame(
  Trait = c('Leaf thickness', 'Leaf area', 'Leaf dry matter content', 'Specific leaf area', 'Wood density'),
  Abbreviation = c('LT', 'LA', 'LDMC', 'SLA', 'WD'),
  Unit = c('$\\mu m$', '$mm^2$', '$mg.g^-1$', '$m^2.kg^-1$', '$g.cm^-3$'),
  Strategy = c('Leaf defense and investment',
        'Leaf resource capture vs.investment',
        'Leaf defense and investment',
        'Leaf resource acquisiton vs. defense',
        'Stem transport, structure and defense')
), caption = 'Functional traits measured. Traits with their abbreviation, standard unit, and associated species strategy for functioning, survival and reproduction [@baraloto_decoupled_2010].', format = 'pandoc')
```

@@

We calculated weighted mean (GWM, where G stands for "guild") and weighted variance (GWV) [GWM and GWV, @violle_let_2007] for each guild in each community. In the quadrat $j$, GWM of trait $t$ was the mean of trait values $t_{i}$ of species $i$ weighted by species relative abundance $p_{i}$:
$$GWM_{j} = \sum_{i=1}^{S} p_{i,j} * t_{i,j} ~~ (1)$$
where $S$ is the number of species in the quadrat $j$. We also calculated the average trait values based on presence-absence values to assess the influence of less abundant and rare species on community trait composition [@cingolani_filtering_2007]. GWV of trait $t$ was calculated as :
$$GWV_{j} = \sum_{i=1}^{S} p_{i,j} * (t_{i,j} - GWM_{j})^2 ~~ (2)$$

We first tested functional compostion differentiation between guilds within each community with a wilcoxson test, to check if the local functional composition from guilds differed. Inside each community we tested the difference between all individuals of each guilds, and condidered the difference to be significative with a p-value inferior to 5%. We then summarized communities significativelly differing in their functional traits between understorey and canpoy guilds for each trait.

Then we devised a null model approach to test for any significant changes in functional composition in response to an environmental gradient. We thus wanted to explore if the local functional composition reflect the influence of abiotic and biotic drivers of community dynamics within guilds. We consequently used subsequent hypothesis:

* ___H1__ Drivers of environmental filtering on local functional composition differ between canopy and understorey tree species._ 
* ___H2__ Strength of drivers of evironmental filtering differ between canopy and understorey tree species._

To accept or reject hypothesis we used two null models :

* __NM1__ randomized trait values among species in the whole plot for each guild (permuting raw in `Trait*Species` matrices of both guilds) [@bernard-verdier_community_2012]. Consequently, NM1 breaks the link between environement and local functional composition while keeping guilds and species spatial structure, thus testing for hypothesis __H1__.
* __NM2__  randomized abundances among species within each guild but kept the observed list of species and trait values unchanged inside guilds [@bernard-verdier_community_2012]. Consequently NM2, breaks the link between environment and species abundances while keeping guilds, community, and species spatial structure, thus testing for hypothesis __H2__.

@@

To characterize the patterns of deviation from the null model, we also calculated standardized effect sizes (SES) as
$$SES = \frac{(I_{obs} - I_{null})}{\sigma _{null}} ~~ (3)$$
where $I_{obs}$ was the observed metric and $I_{null}$ and $\sigma _{null}$ were the mean and standard deviation of the corresponding null distribution. The influence of environment on SES values was investigated by linear regression with biotic and abiotic factors ($SES_{trait} = Slope + Curvature + Wetness + SW + Canopy + BA$). Analyses were conducted on the five functional traits.

@@

```{r CV}
PFT %>% 
  select(-Tree, -Sp, -SpCode, -CE, -SHA, -LTD) %>% 
  melt(variable.name = 'trait') %>% 
  group_by(trait) %>% 
  summarise_each(funs(min(.,na.rm = T), max(., na.rm = T), 
                      mean(.,na.rm = T), sd(.,na.rm = T))) %>% 
  left_join(PFT %>% 
              select(-Sp, -SpCode, -CE, -SHA, -LTD) %>% 
              melt(id.vars = 'Tree', 
                   variable.name = 'trait', 
                   value.name = 'Intra-individuals') %>% 
              group_by(Tree, trait) %>% 
              summarise_each(funs(sd(., na.rm = T)/mean(., na.rm = T))) %>% 
              ungroup() %>% 
              select(-Tree) %>% 
              group_by(trait) %>% 
              summarise_each(funs(mean(., na.rm = T)))) %>% 
  left_join(PFT %>% 
              group_by(Tree, Sp, SpCode, CE) %>% 
              summarise_each(funs(mean)) %>% 
              ungroup() %>% 
              select(-Tree, -SpCode, -CE, -SHA, -LTD) %>% 
              melt(id.vars = 'Sp', 
                   variable.name = 'trait', value.name = 'Intra-species') %>% 
              group_by(Sp, trait) %>% 
              summarise_each(funs(sd(., na.rm = T)/mean(., na.rm = T))) %>% 
              ungroup() %>% 
              select(-Sp) %>% 
              group_by(trait) %>% 
              summarise_each(funs(mean(., na.rm = T)))) %>% 
  left_join(PFT %>% 
              group_by(Tree, Sp, SpCode, CE) %>% 
              summarise_each(funs(mean)) %>% 
              ungroup() %>% 
              select(-Tree, -CE) %>% 
              group_by(Sp, SpCode) %>% 
              summarise_each(funs(mean)) %>% 
              ungroup() %>% 
              select(-Sp, -SpCode, -SHA, -LTD) %>% 
              melt(variable.name = 'trait', value.name = 'Inter-species') %>% 
              group_by(trait) %>% 
              summarise_each(funs(sd(., na.rm = T)/mean(., na.rm = T)))) %>% 
  kable(digits = 2, caption = 'Intra-individual, intra-specific and inter-specific component of trait variation. The overall range, mean and standard deviation are provided for each trait. The remaining columns represent the coefficient of variation of trait values among leaves per individuals (intra-individual), among trees per species (intra-specific), and among trees of different species. See table \\@ref(tab:Traits) for abbreviation.', format = 'pandoc')
```

The first axis of the principal component  analysis (PCA) of mean species trait values was basically related to wood density and structural leaf trait variations (LT, LA, LDMC), thus opposing species with thick, big leaves and lightwood to species with dense leaves and hardwood. The second axis mostly opposed SLA to LDMC, and to a less extent to wood density. However, wood density was orthogonal to SLA in the plan 1-2 of the PCA indicating independence between those two traits. Finally, hypervolume overlap between the two guilds was not significantly lower than null expectations ($p = 0.115$, see Supplementary materials [S2: Hypervolumes of functional guilds]). Still the overlap was pretty low suggesting a trend of functional differentiation between the two guilds.

@@

Except for Specific Leaf Area (SLA), most of the traits were not significantly different between guilds in the majority of communities (only 20 to 41% significative for WD, LDMC, LA and LT against 79% for SLA, see figure \@ref(fig:percsign) ). But regarding guilds weighted means (GWM), understorey and canopy guilds strongly diverged among communities (see figure \@ref(fig:CWMboxplot) ). Canopy guilds have higher SLA, WD and LA but lower LDMC and LT than understorey guilds. In other words, canopy guilds have larger, thinner and lighter leaves with increased SLA and denser woods, whereas understorey guilds have smaller, thicker, and more dense leaves with lower SLA and lighter woods.

```{r percsign, fig.cap="Percentage of communities with indivdual traits significantly different between guilds (for canopy (C) and understorey (U) guilds). Test realized with an unpaired wilcoxon two sided test, significance was considered for p-values under the level $\\alpha = 0.05$ .See table \\@ref(tab:Traits) for abbreviations.", fig.height=4}
L %>% 
  mutate(community = row.names(.)) %>% 
    filter(!(community %in% c("C105", "C11",  "C26",  "C5" ))) %>% 
  melt(id.vars = "community", 
       variable.name = "SpCode",
       value.name = "abundance") %>% 
  filter(abundance > 0) %>% 
  tidyr::uncount(abundance) %>% 
  left_join(Q) %>% 
  select(- SpCode, -LES, -WES) %>% 
  melt(id.vars = c("community", "strata"),
       variable.name = "trait") %>% 
  group_by(community, trait) %>% 
  do(wilcoxon = try(wilcox.test(value~strata, data=.)$p.value)) %>% 
  filter(!inherits(wilcoxon, "try-error")) %>% 
  mutate(wilcoxon = unlist(wilcoxon)) %>% 
  group_by(trait) %>% 
  mutate(test_0.5 = wilcoxon < 0.05) %>%
  dplyr::summarise(com_sign = sum(test_0.5)/n()*100) %>%
  ggplot(aes(trait, com_sign)) +
  geom_bar(stat = "identity") +
  xlab("Functional trait") +
  ylab("Percentage of significant communities")
```

```{r GWMH1a, fig.cap="Number of communities significantly (p-value<0.05) different from null expectations with null model NM1 for each trait of GWMs, for canopy (C) and understorey (U) guilds, in abundance. See table \\@ref(tab:Traits) for abbreviations.", fig.height=4, include = F}
GWM <- L %>%
  t() %>%
  data.frame() %>%
  mutate(SpCode = row.names(.)) %>%
  left_join(Q) %>%
  select(-SpCode) %>%
  melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
       variable.name = 'community', value.name = 'abundance') %>%
  group_by(strata, community) %>%
  summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
# nQ <- array(rep(as.matrix(Q), n),
#             c(dim(Q), n),
#             list(row.names(Q), names(Q), 1:n))
# nQ <- array(apply(nQ, 3, function(x){
#   x[,'strata'] <- sample(x[,'strata'])
#   return(x)}), c(dim(Q), n),
#   list(row.names(Q), names(Q), 1:n))
# cl <- makeCluster(cores)
# clusterExport(cl, list('L', 'nQ'))
# nGWM <- parApply(cl, nQ, 3, function(x){
#   library(dplyr)
#   library(reshape2)
#   L %>%
#     t() %>%
#     data.frame() %>%
#     mutate(SpCode = row.names(.)) %>%
#     left_join(data.frame(x)) %>%
#     select(-SpCode) %>%
#     melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
#          variable.name = 'community', value.name = 'abundance') %>%
#     group_by(strata, community) %>%
#     mutate_each(funs(as.character), -abundance) %>%
#     mutate_each(funs(as.numeric), -abundance) %>%
#     summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance)
# })
# stopCluster(cl)
# nGWM <- array(unlist(lapply(nGWM, function(x) as.matrix(x[3:9]))),
#               c(dim(GWM[3:9]), n),
#               list(row.names(GWM), names(GWM[3:9]), 1:n))
# rm(cl, nQ)
# save(nGWM, file = file.path(path, 'nGWM_NM1.Rdata'))
load(file.path(path, 'nGWM_NM1.Rdata'))
GWM_pval <- GWM
GWM_pval[3:9] <- apply(
  array(c(as.matrix(GWM[3:9]), nGWM), c(dim(GWM[3:9]), n+1)), 
  1:2, function(x) rank(x)[1]/(n+1))
```

```{r CWMboxplot, fig.cap="Boxplot of guilds weighted means (GWM) among communties for each traits (for canopy (C) and understorey (U) guilds). See table \\@ref(tab:Traits) for abbreviations.", fig.height=4, include = T}
GWM %>% 
  select(-community, -LES, -WES) %>% 
  melt(id.vars = "strata", 
       variable.name = "trait",
       value.name = "GWM") %>% 
  ggplot(aes(x = strata, y = GWM, fill = strata)) +
  geom_boxplot() +
  facet_wrap(~trait, scales = "free",nrow = 1) +
  xlab("Guild") + scale_fill_discrete(guide = "none")
```

```{r GWVH1b, fig.cap="Number of communities significantly (p-value<0.05) different from null expectations with null model NM3 for each trait of GWVs, for canopy (C) and understorey (U) guilds, in abundance. See table \\@ref(tab:Traits) for abbreviations.", fig.height=4, include = F}
GWV <- L %>%
  t() %>%
  data.frame() %>%
  filter(sum(.) > 2) %>%
  mutate(SpCode = row.names(.)) %>%
  left_join(Q) %>%
  select(-SpCode) %>%
  melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
       variable.name = 'community', value.name = 'abundance') %>%
  melt(id.vars = c('strata', 'community', 'abundance'),
       variable.name = 'trait', value.name = 'trait_value') %>%
  left_join(melt(GWM, id.vars = c('strata', 'community'),
                 variable.name = 'trait', value.name = 'GWM')) %>%
  group_by(strata, community, trait) %>%
  mutate(GWV = (trait_value - GWM)^2) %>%
  select(-trait_value, -GWM) %>%
  summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) %>%
  dcast(strata + community ~ trait)
# nL <- array(replicate(n, t(apply(L, 1, function(x){
#   x[Q[which(Q$strata == "C"), "SpCode"]][x[Q[which(Q$strata == "C"), "SpCode"]] != 0] <- sample(x[Q[which(Q$strata == "C"), "SpCode"]][x[Q[which(Q$strata == "C"), "SpCode"]] != 0])
#   x[Q[which(Q$strata == "U"), "SpCode"]][x[Q[which(Q$strata == "U"), "SpCode"]] != 0] <- sample(x[Q[which(Q$strata == "U"), "SpCode"]][x[Q[which(Q$strata == "U"), "SpCode"]] != 0])
#   return(x)
#   }))), c(dim(L), n), list(row.names(L), names(L), 1:n))
# cl <- makeCluster(cores)
# clusterExport(cl, list('nL', 'Q', 'GWM'))
# nGWV <- parApply(cl, nL, 3, function(x){
#   library(dplyr)
#   library(reshape2)
#   x %>%
#     t() %>%
#     data.frame() %>%
#     filter(sum(.) > 2) %>%
#     mutate(SpCode = row.names(.)) %>%
#     left_join(Q) %>%
#     select(-SpCode) %>%
#     melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
#          variable.name = 'community', value.name = 'abundance') %>%
#     melt(id.vars = c('strata', 'community', 'abundance'),
#          variable.name = 'trait', value.name = 'trait_value') %>%
#     left_join(melt(GWM, id.vars = c('strata', 'community'),
#                    variable.name = 'trait', value.name = 'GWM')) %>%
#     group_by(strata, community, trait) %>%
#     mutate(GWV = (trait_value - GWM)^2) %>%
#     select(-trait_value, -GWM) %>%
#     summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) %>%
#     dcast(strata + community ~ trait)
# })
# stopCluster(cl)
# nGWV <- array(unlist(lapply(nGWV, function(x) as.matrix(x[3:9]))),
#               c(dim(GWV[3:9]), n),
#               list(row.names(GWV), names(GWV[3:9]), 1:n))
# rm(cl, nL)
# save(nGWV, file = file.path(path, 'nGWV_NM3.Rdata'))
load(file.path(path, 'nGWV_NM3.Rdata'))
GWV_pval <- GWV
GWV_pval[3:9] <- apply(
  array(c(as.matrix(GWV[3:9]), nGWV), c(dim(GWV[3:9]), n+1)), 
  1:2, function(x) rank(x)[1]/(n+1))
GWV_pval %>% 
  select(-community, -LES, -WES) %>% 
  group_by(strata) %>% 
  mutate_all(function(x) findInterval(x, c(0.05, 0.95)) != 1) %>% 
  melt(id.vars = "strata", variable.name = "trait") %>% 
  filter(value == TRUE) %>% 
  ggplot(aes(x = factor(trait), 
             fill = factor(strata))) +
  geom_bar(position = "dodge") +
  scale_fill_discrete("Guild") +
  xlab("Functional trait") +
  ylab("Number of communities") +
  ylim(c(0,dim(L)[1]))
```

```{r CWVboxplot, fig.cap="Caption", fig.height=8, include = F}
GWV %>% 
  select(-community, -LES, -WES) %>% 
  melt(id.vars = "strata", 
       variable.name = "trait",
       value.name = "GWV") %>% 
  ggplot(aes(x = strata, y = GWV, fill = strata)) +
  geom_boxplot() +
  facet_wrap(~trait, scales = "free") +
  ggpubr::stat_compare_means(aes(label = paste("Wilcoxon, p",
                                               ..p.format..,
                                               ..p.signif..)))
```

@@

The guild weighted trait means (GWM) varied weakly but significantly along the environmental gradients for both canopy (C) and understorey (U) guilds (see table \@ref(tab:GWMH2a)). Functional traits of canopy and understorey tree species were both influenced by curvature, but not for the same traits (see figure \@ref(fig:WD)), and two major differences were observed between the two guilds. The effect of south-west orientation on functional composition was observed only for canopy tree species, while the effect of canopy height and wetness affected mainly understorey guild composition.

Focusing on canopy tree species, we found an increase in wood density under south-west orientation. We also observed a decrease in leaf area under the south-west orientation. Leaf area decreased with curvature, while wood density and LDMC increased with curvature. Finally, heigh canopy was also related to an increased leaf area.

For understorey tree species, SLA significantly decreased and and leaf thickness significantly increased with curvature. SLA significantly increased with canopy height, while leaf thickness decreased with canopy height. Finally, SLA and wood density significantly decreased and and leaf area and thickness significantly increased with wetness.
